Getting mortality from atlantisom: backing out F and M from Z

### 2020-01-13 TestCalcFandM.Rmd
initial commit testing calc F and calc M

---
title: "Getting mortality from atlantisom: testing calc_Z function II"
author: "Sarah Gaichas and Christine Stawitz"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

@@

This page documents initial testing of the atlantisom package in development at https://github.com/r4atlantis/atlantisom using three different [Atlantis](https://research.csiro.au/atlantis/) output datasets. Development of atlantisom began at the [2015 Atlantis Summit](https://research.csiro.au/atlantis/atlantis-summit/) in Honolulu, Hawaii, USA. 
On this page we demonstrate use of atlantisom on both California Current (CCA) and Norwegian Barents Sea Atlantis (NOBA) output files to test function `calc_Z` calculating total mortality. This is used both to derive "natural mortality" for comparison with stock assessments and to split age classes into true ages using the function `calc_stage2age`.

NOBA has output files in true ages that we can compare results with. These files are in the folder NOBAwithAnnAgeOutput (not kept on github due to large filesize).

We can also compare both models with the output[scenario.name]Mort.txt file, which reports annual M and F. The user's manual notes: 
>This file is currently only useful for looking at relative M vs F values for a species, as itdoes not give accurate mortalities

So we can assume that the scale of Z from this file will not be the same as our calculation, but it should be useful for splitting out F to get M, and for comparing overall patterns.

We will first use CCA here because the contrasting F scenario has been implemented. We will test with both sardines (M needed, but not age splitting) and hake (M and age splitting needed).

@@

initCCA <- TRUE
initNEUS <- FALSE
initNOBA <- FALSE

@@

if(initCCA) source(here("config/CC2Config.R"))

if(initNEUS) source(here("config/NEUSConfig.R"))

@@

<!-- not sure we need catch bio for this
```{r loadcatchb}

truecatchbio <- load_catch(d.name, file_catch = catch.file, fgs = funct.groups)

truecatchbio_ss <- truecatchbio[truecatchbio$species == species_ss,]
# note: time output of this file is in days
# what model timestep is this output matching?
# rescale catch in biomass time from days to years
truecatchbio_ss$time <- as.integer(truecatchbio_ss$time/365)

## Test calc Z function

>>>>

---
title: "Getting mortality from atlantisom: backing out F and M from Z"
author: "Sarah Gaichas and Christine Stawitz"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

@@

This page documents initial testing of the atlantisom package in development at https://github.com/r4atlantis/atlantisom using three different [Atlantis](https://research.csiro.au/atlantis/) output datasets. Development of atlantisom began at the [2015 Atlantis Summit](https://research.csiro.au/atlantis/atlantis-summit/) in Honolulu, Hawaii, USA. 
On this page we demonstrate use of atlantisom on both California Current (CCA) and Norwegian Barents Sea Atlantis (NOBA) output files to test function new functions that use `calc_Z` calculating total mortality to back out F and M. These are used initially for comparison with stock assessment model fishing mortality outputs, and to parameterize stock assessment natural mortality.

[Previous comparisons](https://sgaichas.github.io/poseidon-dev/TestCalcZ2.html) of estimated total mortality with the output[scenario.name]Mort.txt file demonstrated what the Atlantis manual already tels us: 
>This file is currently only useful for looking at relative M vs F values for a species, as itdoes not give accurate mortalities

We can assumed that the scale of Z from this file will not be the same as our calculation, but since both F and M are rescaled, even from input F values, we need to back out the realized F to subtract from this Z to get M. This approach seems simpler than trying to estimate M directly using predation.

We will first use CCA here because the contrasting F scenario has been implemented. We will test with sardines only first.

@@

initCCA <- TRUE
initNOBA <- FALSE

@@

if(initCCA) source(here("config/CC3Config.R"))

@@

## Estimate total mortality with calc Z function
